ISF Grant

### 2017-10-08 article.tex
Model 3
Model done.
to do: references + formal definitions

The standard abstract model is consider in this work. The system consist of set of asynchronous unreliable processes $P=\{p_1,\ldots,p_n\}$ that communicate by applying operations on shared \emph{base objects} that support atomic operations such as reads, writes, and read-modify-write primitives.

We assume that a once a process crash, the system knows at which line of the code the fail happened. However, there is no indication whether the instruction at that line tool affect before the crash. Upon recovering, the controller moves the process to a recovery code according to the line where the crash occur. By running this code, the process is aware that a crash happened, as well as at what line of the code was it.

@@

The correctness properties variants discussed earlier does not need to consider primitives, as any atomic operation satisfies any of the properties. Unlike it, RR-linearizability is not satisfies by primitives since the respond of the primitive is lost in case of a failure. Therefore, there is a need to supply each primitive with an implementation as well as a recovery code. In the following section we do so for well known primitives, as well as presenting impossibility result for others. We assume a failure model of a full-system crash, and address the wait-free progress property.
The fact that RR-linearizabilty allows us to swift the linearization point of an operation to after the crash is used for the recover after a primitive failure.

@@

\paragraph*{Test-and-Set}
We now show that there is no wait-free RR-linearizability implementation for TAS for two processes, assuming the system supply only read/write and TAS primitives. We give a high level description for the proof.

>>>>

The standard abstract model is consider in this work. The system consist of a set of asynchronous unreliable processes $P=\{p_1,\ldots,p_n\}$ that communicate by applying operations on shared \emph{base objects} that support atomic operations such as reads, writes, and read-modify-write primitives. All base objects are non-volatile, i.e, in case of a crash the values of the base objects is preserved. In order to implement a more complex object, such as stack and counter, each process is supplied with access procedures that simulate each operation on the implemented object using operations on base objects.

The failure model considered in this paper is individual. Each process may crash, recover, and then continue executing, at any point of the execution. Once a process crash, we assume that the system control knows at which line of the process code the failure occurs, or more precisely, at the course of which atomic operation the process failed. However, there is no induction whether the atomic operation took affect or not before the crash. Each process is also equipped with a recovery code for any possible failure, that is, for any point of the execution it specifies the recovery code in case of a crash at that point. Upon recovering, the system control moves the process to the beginning of the appropriate recovery code. In this sense, by running the recovery code, the process is aware of the crash. At the end of the recovery code, the process goes back to the main code, to the location specified by the recovery code. The above model can be implemented by storing the program counter at the non-volatile memory. In this case, after each operation on a base object, there should be an operation that updates the value of the program counter. This can be done by the system control, and the programmer need not be aware of it. Despite doubling the number of accesses to the main memory, this approach reduce the complexity of recovering in case of a crash.

Real world system presents more complications to the abstract model described above. In most modern system, each process has its own local volatile cache, which is used to store variables accessed by the process. Moreover, in order to improve performance in such systems, writes do not go directly to the shared memory, but rather stored in a buffer at the process's cache, while the system commit the writes later on according to the protocol and model it satisfies. Real system thus introduce reordering of instructions, as well as the possibility of losing large amount of data in case of a failure, which makes the recovery procedure harder and more complicate. However, the abstract model can be simulated in many of those architectures using special instructions as memory fences and avoid using the cache. The cost of such simulation nonetheless, might be very expensive in terms of running time.
The results presented in this paper implies that even in the simple abstract model the recovery procedure is not trivial. For future work, more complicated and realistic models needs to be studied.

@@

The correctness properties variants discussed earlier does not need to consider primitives, as any atomic operation satisfies any of the properties. Unlike it, RR-linearizability is not satisfies by primitives since the respond of the primitive is lost in case of a failure. Therefore, there is a need to supply each primitive with an implementation as well as a recovery code. A system which supply recoverable primitives can be used to implement any object in a recoverable way - in case of a crash the process will simply recover the last atomic operation it executed before the crash. In case the operation is aborted, the process will reissue it. After that, it can continue and execute the rest of its code safely. This observation focus our attention to implementation of primitives in a recoverable manner.

In the following section we present RR-liniarize implementations for well known primitives, as well as presenting impossibility result for others. We focus our attention on bounded wait-free implementations, that is, the number of steps a process takes when executing the recovery code in the absence of a failure is finite and bounded by a known constant (may be a function of n, the number of processes in the system), regardless of the other processes steps and the failures the process experienced before. In addition, we would like the recovery code to use a finite number of variables.
The fact that RR-linearizabilty allows us to swift the linearization point of an operation to after the crash is used to recover after a primitive failure.

@@

\paragraph*{Test-and-Set}

Test-and-Set (TAS) is a known object supported as a primitive in many modern architectures. A TAS object contains a binary value which initialized to 0, and support a single atomic operation TAS which set the value of the object to 1 and returns the old value.

We now show that there is no bounded wait-free RR-linearizability implementation for TAS for two processes, assuming the system supply only read/write and TAS primitives. We give a high level description for the proof. The construction also implies that any wait-free implementation need to use an unbounded number of TAS objects.
